\mythesistitle

### 2011-11-23 src/synopsis.tex
all sections except sec 4 onwards are done. rest will be done in less than an hour

\tnote{testing 1 2 3...}
\tnote[bogus1]{testing 1 2 3...}
\tnote[bogus2]{testing 1 2 3...}
\tnote[GTC]{testing 1 2 3...}
\tnote[bogus4]{testing 1 2 3...}
\tnote[TH]{testing 1 2 3...}

@@

A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
  \Sc\ and \Lu\ enroll at the
% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
% enroll at the
{\WSI} for a liberal arts programme.  As part
of their semester thesis, they pick a body of work to study and form
the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
{\TTT}. \tnote[TH]{put bib entries for these works!} A student
will be in at least one study group and may be in more than one. For
instance, as will be seen later, {\Fr} studies both {\LLL} and {\TTT}
while \Wo\ studies only \BBB.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{rcl}
    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
    $n $&$=$&$ |U| = 11$\\
    $m $&$=$&$ |\cF| = 4$      
  \end{tabular}
  \caption{Students and study groups in \WSI}
  \label{tab:wsigroups}
\end{table}

@@

An intuitive approach to this problem would be to first find the paths
that each study group decides to inhabit and then refine the
allocation to individual students. A feasible allocation of exclusive
routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
and the students' allocation of apartments that obeys this route
allocation is shown in Fig~\ref{fig:streetmappathpeople}. How this is
algorithmically computed is the focus of this thesis.

\begin{figure}[t]%[htbp]
 \centering
  \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
  \caption{{\residenceblock} street map}
  \label{fig:streetmap}
\end{figure}

\begin{figure}[t]%[htbp]
  \centering
  \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
  \caption{{\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
  \label{fig:streetmappath}
\end{figure}

\begin{figure}[t] % [h]
  \centering
  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
  \caption{Individual allocation of apartments to students in
    {\residenceblock} that meets the requirements stated before.}
  \label{fig:streetmappathpeople}
\end{figure}

@@

\tnote{add a note on the organization of the doc}
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
% Section~\ref{sec:feasible} documents the
% characterization of a feasible path labeling and finally,
% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
% find the tree path labeling of a given set system from a given
% $k$-subdivided tree.

Section~\ref{sec:motivation} gives a brief survey of COP and
optimization problems related to it.  Section~\ref{sec:results}
presents a summary of our results on the extention of COP namely, the
tree path labeling problem.

\section{Motivation and background}
\label{sec:motivation}

As seen earlier, the interval assignment problem (illustrated as the
course scheduling problem in Section~\ref{sec:problem}), is a special
case of the problem we address in this thesis, namely the tree path
labeling problem (illustrated as the study group accommodation
problem). Interval assignment and COP are equivalent
problems. In this section we will see some of the results that
exists in the literature today towards solving the COP problem and
optimization problems surrounding it.

@@

\begin{figure}[t] %[htbp]
  \centering
  % \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
  \begin{tabular}[h]{l|lcccl}
    $M_1$: & $M_1'$: &&&& $M_2$:\\
    &&&&&\\
    \begin{tabular}[h]{llll}
      $c_1$ & $c_2$ &$c_3$ &$c_4$\\
      &&&\\
      \un & 0   & \un & 0\\
      0   & \un & 0   & \un \\
      \un & 0   & 0   & \un
    \end{tabular}
    &
    \begin{tabular}[h]{llll}
      $c_3$ &$c_1$ &$c_4$& $c_2$\\
      &&&\\
      \un & \un & 0 & 0\\
      0 & 0 & \un & \un \\
      0 & \un & \un & 0
    \end{tabular}
    &&&&
    \begin{tabular}[h]{llll}
      $d_1$ & $d_2$ &$d_3$ &$d_4$\\
      % &&&\\
      &&&\\
      \un & \un & 0 & 0\\
      0 & \un & \un & 0 \\
      0 & \un & 0 & \un 
    \end{tabular}
  \end{tabular}
  \caption{Matrices with and without COP. $M_1$ has COP because by
    permuting its columns, $c_1$-$c_4$, one can obtain $M_1'$ where
    the {\un}s in each row are consecutive. $M_2$, however, does not
    have COP since no permutation of its columns, $d_1$-$d_4$, will
    arrange {\un}s in each row consecutively \cite{d08phd}.}
  \label{fig:cop-matrix}
\end{figure}

@@

MOVE TO THESIS?
-- sect 4.1 in \cite{d08phd} has many results surveyed. hardness
results, approx. results. results are usually for a class of matrices
$(a,b)$ where number columns and rows are restriced to $a$ and $b$
\tnote{check if it is the other way around}.\\
-- problem of flipping at most $k$ entries of $M$ to make it attain
COP. this is NP complete \cite{b75-phd}
END MOVE TO THESIS?

\tnote[TH]{cite:lb62 - interval graphs are at-free chordal}

\cite{at72} showed that a matrix that does not have COP have
certain substructures that prevent it from having
COP. Tucker classified these forbidden substructures into five classes of
submatrices. This result is presented in the context of convex
bipartite graphs which \cite{at72} proved to be
AT-free\tnote[TH]{ check this up. give details. - doms'}. Convex bipartite
graph have vertex incidence adjacency matrices is
equivalent to matrices with COP. A binary matrix is converted into a
bipartite graph as follows. The set of rows and the set of columns
form the two partitions of the graph. Each row node is adjacent to
those nodes that represent the columns that have ones in the
corresponding row. \cite{at72} proves that this bipartite graph has no
asteroidal triple if and only if the matrix has COP and goes on to
identify the forbidden substructures for these bipartite graphs. The
matrices corresponding to these substructures are the forbidden
submatrices. Once a matrix has been detected to not have COP (using
any of the COT algorithms mentioned earlier), it is naturally of
interest to find out the smallest (in terms of number of rows, columns
and number of entries that are ones) forbidden
substructure. \cite{d08phd} discusses an exact algorithm with time
complexity \annote{parameterized by the number of ones in the
  matrix}{check this.}.

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete\tnote{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hg02} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem\tnote[TH]{elaborate - what are
  the results?}.

Another problem is to find the minumum number of entries in the matrix
that can be flipped to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to ones so that the
resulting matrix has COP. As mentioned earlier, this problem had been
known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
using a reduction to the longest path problem, \tnote{or is it a
  survey of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$ rows is NP complete. \tnote{how is this
  different from booth's 75 result??}  \tnote[TH]{where should this
  go? cite|tz04 (approx submatrix with COP sparse matrices)}

\cite{jkckv04} discusses the use of matrices with almost-COP in the
storage of very large databases.  The problem is that of $k$-run
reordering of binary matrices which they prove to be NP hard using a
reduction from the hamiltonian path problem.  A connection of COP
problem to the travelling salesman problem is also
introduced. \tnote[TH]{what does this mean? -- COP can be used as a
  tool to reorder $0.5T \le runs(M) \le T$}

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{k77}. \tnote[TH]{what is the reduction?}

\tnote[TH]{other problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

{\bf Motivation: } A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

How the problem of COT is related to the constraint statisfaction
problem of interval assignment to sets in a system can be seen as
follows. Every column of the binary matrix can be converted into a set
of integers which are the indices of rows with {\un}s in that column. It
is apparent that if the matrix has COP, then constructing such sets
after applying the COP permutation to the matrix will result in all
sets with consecutive integers. In other words, after application of
COP reordering, the sets are intervals. Indeed the problem now becomes
finding interval assignments to a given set system such that there
exists a permutation of the universe (set of row indices) which
converts each set to its assigned interval. The result in
\cite{nsnrs09} characterizes interval assignments to the sets which
can be obtained from a single permutation of the rows.  They show that
for each set, the cardinality of the interval assigned to it must be
same as the cardinality of the set, and the intersection cardinality
of any two sets must be same as the intersection cardinality of the
corresponding intervals.  While this is obviously a necessary
condition, this result shows this is also sufficient.  \cite{nsnrs09}
calls such an interval assignment an Intersection Cardinality
Preserving Interval Assignment (ICPIA).  This paper generalizes the
idea from \cite{wlh02} of decomposing a given binary matrix into prime
matrices for COT and describes an algorithm to test if an ICPIA exists
for a given set system.

@@

Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}.

%\item 
In the Section 4, \annote{we initiate an exploration of}{WTF?}
finding feasible path labeling of set systems in a special kind of
tree which we call the $k$-subdivided star.  \annote{This question
  is an attempt to generalize}{REWORD.} the problem of testing if a
matrix has the consecutive ones property.  However, we restrict the
hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
elements.  \annote{In spite of this restricted case we consider, we
  believe that our results are of significant interest in
  understanding the nature of Graph Isomorphism}{REWORD.} which is
polynomial time solvable in interval graphs and is hard on path
graphs.
%\end{enumerate}

@@

\section{Conclusion}
\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[TH]{peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the

>>>>

% \tnote{testing 1 2 3...}
% \tnote[bogus1]{testing 1 2 3...}
% \tnote[bogus2]{testing 1 2 3...}
% \tnote[GTC]{testing 1 2 3...}
% \tnote[bogus4]{testing 1 2 3...}
% \tnote[TH]{testing 1 2 3...}

@@

A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
  \Sc\ and \Lu\ enroll at the
% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
% enroll at the
  {\WSI} for a liberal arts programme.  As part of their semester
  thesis, they pick a body of work to study and form the namesake
  study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}\tnote[TH]{put bib
    entries for these works!}. A student will be in at least one study
  group and may be in more than one. For instance, as will be seen
  later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\ studies only
  \BBB.

@@

\def \figtabsize {\footnotesize}

\begin{table}[htbp]
  \centering
{
  \begin{tabular}{rcl}
    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
    $n $&$=$&$ |U| = 11$\\
    $m $&$=$&$ |\cF| = 4$      
  \end{tabular}
}
  \caption{\figtabsize Students and study groups in \WSI}
  \label{tab:wsigroups}
\end{table}

@@

An intuitive approach to this problem would be to first find the paths
that each study group decides to inhabit and then refine the
allocation to individual students. A feasible allocation of exclusive
routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
and the students' allocation of apartments that obeys this route
allocation is also shown\tnote[TH]{in
  Fig~ref::fig:streetmappathpeople}. How this is algorithmically
computed is the focus of this thesis.

%%%%%%%%%%%% THESIS UNCOMMENT %%%%%%%%%%%%
% \begin{figure}[t]%[htbp]
%  \centering
%   \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
%   \caption{{\residenceblock} street map}
%   \label{fig:streetmap}
% \end{figure}

% \begin{figure}[t]%[htbp]
%   \centering
%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
%   \caption{{\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
%   \label{fig:streetmappath}
% \end{figure}
% \begin{figure}[t] % [h]
%   \centering
%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
%   \caption{Individual allocation of apartments to students in
%     {\residenceblock} that meets the requirements stated before.}
%   \label{fig:streetmappathpeople}
% \end{figure}
%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
\begin{figure}[htbp] %[t]%
 \centering
  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
 \label{fig:streetmap}
  \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
\end{figure}
\begin{figure}[htbp] %[t]%
  \centering
  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
  \caption{\figtabsize Individual allocation of apartments to students in
    {\residenceblock} that meets the requirements stated before.}
  \label{fig:streetmappathpeople}
\end{figure}

@@

Section~\ref{sec:background} gives a brief survey of COP and
optimization problems related to it.  Section~\ref{sec:results}
presents a summary of our results on the extention of COP namely, the
tree path labeling problem. \tnote[TH]{update for thesis}

\section{Background}
\label{sec:background}
\subsection{Matrices with COP}
\label{sec:copmatrices}
As seen earlier, the interval assignment problem (illustrated as the
course scheduling problem in Section~\ref{sec:problem}), is a special
case of the problem we address in this thesis, namely the tree path
labeling problem (illustrated as the study group accommodation
problem). Interval assignment and COP are equivalent
problems. In this section we will see some of the results that
exists in the literature today towards solving the COP problem and
optimization problems surrounding it.

@@

\begin{figure}[t] %[htbp]
  \centering
  % \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
{\figtabsize
 \begin{tabular}[h]{l|lcccl}
    $M_1$: & $M_1'$: &&&& $M_2$:\\
    &&&&&\\
    \begin{tabular}[h]{llll}
      $c_1$ & $c_2$ &$c_3$ &$c_4$\\
      &&&\\
      \un & 0   & \un & 0\\
      0   & \un & 0   & \un \\
      \un & 0   & 0   & \un
    \end{tabular}
    &
    \begin{tabular}[h]{llll}
      $c_3$ &$c_1$ &$c_4$& $c_2$\\
      &&&\\
      \un & \un & 0 & 0\\
      0 & 0 & \un & \un \\
      0 & \un & \un & 0
    \end{tabular}
    &&&&
    \begin{tabular}[h]{llll}
      $d_1$ & $d_2$ &$d_3$ &$d_4$\\
      % &&&\\
      &&&\\
      \un & \un & 0 & 0\\
      0 & \un & \un & 0 \\
      0 & \un & 0 & \un 
    \end{tabular}
  \end{tabular}
}

  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP because by
    permuting its columns, $c_1$-$c_4$, one can obtain $M_1'$ where
    the {\un}s in each row are consecutive. $M_2$, however, does not
    have COP since no permutation of its columns, $d_1$-$d_4$, will
    arrange {\un}s in each row consecutively \cite{d08phd}.}

  \label{fig:cop-matrix}
\end{figure}

@@

\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
  surveyed. hardness results, approx. results. results are usually for
  a class of matrices $(a,b)$ where number {\un}s in columns and rows
  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
  entries of $M$ to make it attain COP. this is NP complete
  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
  are AT-free.  describe AT (2) show the close relationship b/w COP
  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
does not have COP have certain substructures that prevent it from
having COP. Tucker classified these forbidden substructures into five
classes of submatrices. This result is presented in the context of
convex bipartite graphs which \cite{at72} proved to be
AT-free\tnote[TH]{ check this up. give details. - doms'}. By
definition, convex bipartite graph have half adjacency matrices that
have COP on either rows or columns (graph is biconvex if it has COP on
both)\cite{d08phd}. A half adjacency matrix is a binary matrix
representing a bipartite graph as follows. The set of rows and the set
of columns form the two partitions of the graph. Each row node is
adjacent to those nodes that represent the columns that have ones in
the corresponding row. \cite{at72} proves that this bipartite graph
has no asteroidal triple if and only if the matrix has COP and goes on
to identify the forbidden substructures for these bipartite
graphs. The matrices corresponding to these substructures are the
forbidden submatrices.

Once a matrix has been detected to not have COP (using any of the COT
algorithms mentioned earlier), it is naturally of interest to find out
the smallest (in terms of number of rows and/or columns and/or number
of entries that are {\un}s) forbidden substructure. \cite{d08phd}
discusses a couple of algorithms which are efficient if the number of
ones in a row is small. This is of significance in the case of sparce
matrices where this number is much lesser than the number of
columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
on number of {\un}s in any column but has at most $\Delta$ {\un}s in
any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
similar problems which deals with inducing COP on a matrix. In {\sc
  Min COS-R (Min COS-C)} the question is to find the minimum number of
rows (columns) that must be deleted to result in a matrix with COP.
In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
maximum number of rows (columns) that induces a submatrix with
COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
a submatrix $M'$ with all columns\tnote{check if b75 deals with COP
  col or COP row. also is it any submatrix with k less than r rows or
  submatrix must have all columns?} but a maximum cardinality subset
of rows such that $M'$ has COP is NP complete. \cite{hg02} corrects an
error of the abridged proof of this reduction as given in \cite{gj79}.
\cite{d08phd} discusses all these problems in detail giving an
extensive survey of the previously existing results which are almost
exhaustively all approximation results and hardness results. Taking
this further, \cite{d08phd} presents new results in the area of
parameterized algorithms for this problem\tnote[TH]{elaborate - what
  are the results?}.

Another problem is to find the minimum number of entries in the matrix
that can be toggled to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to {\un}s so that the
resulting matrix has COP. As mentioned earlier, this problem is
known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
using a reduction to the longest path problem, \tnote{or is it a
  survey of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$ rows is NP complete. \tnote{how is this
  different from booth's 75 result??}  \tnote[TH]{where should this
  go? cite|tz04 (approx submatrix with COP sparse matrices)}

\cite{jkckv04} discusses the use of matrices with almost-COP (instead
of one block of consecutive {\un}s, they have $x$ blocks, or {\em
  runs}, of consecutive {\un}s and $x$ is not too large) in the
storage of very large databases.  The problem is that of reordering of
a binary matrix such that the resulting matrix has at most $k$ runs of
{\un}s. This is proven to be NP hard using a reduction from the
hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
\tnote[TH]{(1) A connection of COP problem to the travelling salesman
  problem is also introduced. what does this mean? -- COP can be used
  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
version of the $k$-run problem, i.e. minimization of number of blocks
of ones is proven to be NP complete by cite:k77}\tnote{are these two
  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

\section{Motivation}
\label{sec:motive}
As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
characterization of consecutive ones property based on the cardinality
properties of the set representations of the columns. This result is
very relevant to this thesis because aside from it simplifying COT to
a great degree, our generalization problem is motivated by their
results.

The result in
\cite{nsnrs09} characterizes interval assignments to the sets which
can be obtained from a single permutation of the rows.  They show that
for each set, the cardinality of the interval assigned to it must be
same as the cardinality of the set, and the intersection cardinality
of any two sets must be same as the intersection cardinality of the
corresponding intervals.  While this is obviously a necessary
condition, this result shows this is also sufficient.  \cite{nsnrs09}
calls such an interval assignment an Intersection Cardinality
Preserving Interval Assignment (ICPIA).  This paper generalizes the
idea from \cite{wlh02} of decomposing a given binary matrix into prime
matrices for COT and describes an algorithm to test if an ICPIA exists
for a given set system.

The tree path labeling problem is a natural generalization of the
interval assignment problem or the COP problem. The problem is defined
as follows -- given a set system $\cF$ from a universe $U$ and a tree
$T$, does there exist a bijection from $U$ to the vertices of $T$ such
that each set in the system maps to a path in $T$.  We refer to this
as the {\em tree path labeling problem} for an input set system,
target tree pair -- $(\cF,T)$. As a special case if the tree $T$ is a
path, the problem becomes the interval assignment problem.  We focus
on the question of generalizing the notion of an ICPIA \cite{nsnrs09}
to characterize feasible path assignments.  We show that for a given
set system $\cF$, a tree $T$, and an assignment of paths from $T$ to
the sets, there is a feasible bijection between $U$ and $V(T)$ if and
only if all intersection cardinalities among any three sets (not
necessarily distinct) is same as the intersection cardinality of the
paths assigned to them and the input passes a filtering algorithm
(described in this paper) successfully.  This characterization gives a
natural data structure that stores all the relevant feasible
bijections between $U$ and $V(T)$. This reduces the search space for
the solution considerably from the universe of all possible bijections
between $U$ and $V(T)$ to only those bijections that maintain the
characterization\tnote[TH]{REWORD}.  Further, the filtering algorithm
is also an efficient algorithm to test if a tree path labeling to the
set system is
feasible.  %This generalizes the result in \cite{nsnrs09}.

% How the problem of COT is related to the constraint statisfaction
% problem of interval assignment to sets in a system can be seen as
% follows. Every column of the binary matrix can be converted into a set
% of integers which are the indices of rows with {\un}s in that column. It
% is apparent that if the matrix has COP, then constructing such sets
% after applying the COP permutation to the matrix will result in all
% sets with consecutive integers. In other words, after application of
% COP reordering, the sets are intervals. Indeed the problem now becomes
% finding interval assignments to a given set system such that there
% exists a permutation of the universe (set of row indices) which
% converts each set to its assigned interval.

@@

We address the computation of TPL\footnote{Tree Path Labeling} for a
given set system and tree in this research. However optimization
problems have not been explored.

Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}, in
other words the interval assignment problem or consecutive ones problem.

%\item 
Aside from checking if a given TPL is feasible, we also attempt to
compute the TPL for a given set system (hypergraph) and tree, if one
exists. For the latter problem, we have an algorithm for general trees
which runs in exponential time. We present a polynomial time
algorithm for a special class of trees called $k$-subdivided trees and
when the hypergraph $\cF$ is such that all hyperedges have at most
$k+2$ elements.  In spite of this restricted case we consider, we
believe that our results are of significant interest in understanding
the nature of Graph Isomorphism which is polynomial time solvable in
interval graphs and is hard on path graphs. The connection of TPL to graph
isomorphism will be made later in the document
%\end{enumerate}

@@

\section{Conclusion}

\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[TH]{peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the
